Independent discovery of a nulling pulsar with unusual sub-pulse drifting properties with the Murchison Widefield Array

### 2022-05-12 paper/referee_report_1/response.md

Dear editor,

We thank the referee for the many detailed comments, which we have tried to address in this revised version. In many instances, the referee asked for a more thorough analysis, particularly surrounding the nulling phenomenon, which -- now that the paper is being submitted to the main journal (instead of Letters) there has been room to include, and we believe the paper is much better for it. In a few cases, however, we have deferred the suggested deeper analysis until the planned follow-up paper, which will benefit from an expanded data set and a wider frequency coverage. Please find our detailed response to the referee's comments below.

Sincerely,

~/Sam

@@

8. It is difficult to determine whether nulling or just weak emission exists in this pulsar only from Figures 2 and 3. The pulse energy distribution is a statistical way to characterize the nulling behavior (Wen et al. 2020, ApJ, 904, 72).

  * We have now included pulse energy distributions for all six observations (see also our response to points 12 & 13 below). From the distributions of the two 20-min observations in which the pulsar was not detected in a periodicity search, we conclude that if there is any weak emission, it must fall far below the MWA's detection threshold. The new text discussing the pulse energy histograms can be found in Section 3.1.

9. Is there any temporal evolution of nulling fraction from follow-up observations and archival MWA data? Is it possible that the constraints on the nulling fraction might be reported in this paper?

  * It still remains the case that only a limited number of archival MWA data sets have been processed and analysed (being those which are presented in this paper). We do expect to get better constraints of the nulling fraction once more data have been processed, but this remains outside the scope of the present initial analysis, and will be presented in the planned follow-up paper.

@@

10. The nulls from a few pulsars does not seem to be random, but show well-defined periodicity (e.g. Herfindal & Rankin 2007, MNRAS, 380, 430; Wang et al. 2021, ApJ, 923, 259). Try to examine whether the periodicity in the transitions between null and burst states is present in this pulsar.

  * We have examined the fluctuation spectra for each of our data sets, but find no evidence of periodicity in the null-burst transitions. To be clear, the observations which showed any discernible low-frequency features at all were 1226062160, 1274143152 and 1283104232. For the latter two, the low-frequency feature occurred in the first (non-DC) frequency bin, which is not surprising, as this only reflects the fact that only a single (long) nulling sequence was present in both. Similarly, the low-frequency feature in 1226062160's spectrum occurs in the second (non-DC) bin, corresponding to the presence of two main (long) null sequences. In all cases, the ratio of observation length to null length is too small to make any assertions about whether the observed null lengths are typical, or whether they are periodic or quasi-periodic. This will be explored in the future once more (and longer) follow-up observations have been made.

11. The pulsar switches between null and burst states with short and long timescales, which provides an important clue to scrutinize various nulling mechanism models. The null and burst length histograms are required.

  * We completely agree that a deeper study of the state transitions has great potential to shed light on the nulling mechanism. However, the working definition of nulls used in this work makes the suggested null and burst length analysis problematic, because no distinction is made between "true" nulls and pseudo-nulls. Consequently, long burst sequences that are likely continuous under a stricter definition of nulls are, in this analysis, frequently interrupted by (likely) pseudo-nulls (as can be clearly seen in the new Figure 2). This effectively alters the null and burst length histograms dramatically. Therefore, we feel that histograms of the null/burst lengths under the current definition would not contain as much useful information about the nulling mechanism as hoped. A deeper analysis is certainly required, but, we feel, outside the scope of the present paper and will be included in the follow-up paper.

### 2022-05-12 paper/main.tex
Added in Ramesh's feedback

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%%
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands.
%%   tighten      : Makes baselineskip slightly smaller, only works with
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much
%%                  longer author list but may be desirable for short
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\psr}{J0026$-$1956}
\newcommand{\psrdm}{20.869}
\newcommand{\psrdmerr}{0.005}
\newcommand{\psrnick}{J0036$-$1033}
\newcommand{\dmunit}{pc$\,$cm$^{-3}$}
\newcommand{\Sd}{S$_\rm{d}$}
\newcommand{\nullfraction}{78\%}
\newcommand{\ExB}{\ensuremath{{\bf E}\times{\bf B}}}

@@

{\bf
Pulsars that both show complicated drifting behavior and are bright enough for single pulse analysis, are relatively rare.
}
Even a cursory glance at the drift bands of \psr{} show that its drifting behavior is indeed complex and bright, making it a welcome addition to this category of pulsar.
{\bf 
\psr{} is also found to contain null sequences, a phenomenon which is known to be connected to sub-pulse drifting, originally noticed by \citep{Lyne1983}, and further attested by recent studies of PSRs J1727$-$2739 \citep{Wen2016}, J1822$-$2256 \citep{Janagal2022}, and the Vela pulsar \citep{Wen2020}.
}

@@

Follow-up of the initial MWA detection with the full 80 minutes, as well as in other archival MWA observations, revealed that the pulsar has a very large nulling fraction ($\sim$~\nullfraction{})\footnote{The nulling fraction estimate comes only from analysis of the MWA observations.}, with some nulling sequences {\bf (excluding occasional intermittent pulses) potentially exceeding $20\,$min (i.e. the length of the first observation on MJD 59002, which contained \textit{only} a few intermittent pulses, as shown in Fig. \ref{fig:maxima})}.
A full analysis of the nulling properties is presented in Section \S\ref{sec:nulling} below.
Despite the fact that \psr{} was serendipitously discovered in the first ten minutes, it may well be the harbinger for a population of relatively bright, intermittent pulsars which the SMART survey, with its $80$-minute dwell times, is well-suited to detect.
A similar strategy is employed in the Low Frequency Array (LOFAR) Tied-Array All-Sky Survey \citep[LOTAAS][]{Sanidas2019} and the low latitude ($|b| \le 5^\circ$) segment of the High Time Resolution Universe survey \citep[HTRU][]{Keith2010}, which use dwell times of approximately one hour.

@@

The other {\bf five observations} were taken when the MWA was in its extended configuration, giving a tied-array beam size of $\sim3^\prime$ at $155\,$MHz.
Localization within {\bf the four observations in which the pulsar was bright enough} followed the identical procedure as used in \citet{Swainston2021}, i.e.\ gridding the candidate positions with overlapping tied-array beams and using signal-to-noise and the estimated beam shape to produce a likelihood map of the source position.
It is known that the ionosphere is capable of shifting the apparent position of sources by an appreciable fraction ($\sim10$-$30\%$) of the tied-array beam, and that these offsets are not necessarily corrected by the calibration process when the calibration observations are separated from the target observation, both spatially and temporally \citetext{Swainston et al., submitted}.
Nevertheless, the localization afforded by the extended array observations was sufficiently precise (a few arcminutes) to warrant follow-up with a high-resolution imaging telescope.
Accordingly, images made with Band 3 ($300$-$500\,$MHz) data from the upgraded Giant Metrewave Radio Telescope \citep[uGMRT;][]{Gupta2017}, obtained under Director's Discretionary Time (DDTC185), enabled us to confirm the pulsar's position to an accuracy of $\sim1^\prime$.
The right ascension and declination of the final best position, consistent with both the uGMRT imaging and the MWA extended configuration observations, are 00h26m37s and $-19^\circ$56'28''.

@@

\begin{figure*}[t]
    \includegraphics[width=0.32\textwidth]{1226062160_nullhist.png}
    \includegraphics[width=0.32\textwidth]{1274143152_nullhist.png}
    \includegraphics[width=0.32\textwidth]{1275094456_nullhist.png} \\
    \includegraphics[width=0.32\textwidth]{1275172216_nullhist.png}
    \includegraphics[width=0.32\textwidth]{1275178816_nullhist.png}
    \includegraphics[width=0.32\textwidth]{1283104232_nullhist.png}
    \caption{{\bf The pulse energy histograms for each of the six observations, for matched-size on- and off-pulse regions of the pulse stacks. The ObsIDs are given above each plot. The on-pulse histograms include all pulses, regardless of whether they were classified as nulls by the method described in the text.}}
    \label{fig:nullhist}
\end{figure*}

>>>>

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be invoked like this:
%%
%% \documentclass[arguments]{aastex631}
%%
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts
%%                but now that abstracts are required this is obsolete for
%%                AAS Journals. Authors might need it for other reasons. DO NOT
%%                use \begin{abstract} and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can invoke to allow other stylistic
%% actions. The available options are:
%%
%%   astrosymb    : Loads Astrosymb font and define \astrocommands.
%%   tighten      : Makes baselineskip slightly smaller, only works with
%%                  the twocolumn substyle.
%%   times        : uses times font instead of the default
%%   linenumbers  : turn on lineno package.
%%   trackchanges : required to see the revision mark up and print its output
%%   longauthor   : Do not use the more compressed footnote style (default) for
%%                  the author/collaboration/affiliations. Instead print all
%%                  affiliation information after each name. Creates a much
%%                  longer author list but may be desirable for short
%%                  author papers.
%% twocolappendix : make 2 column appendix.
%%   anonymous    : Do not show the authors, affiliations and acknowledgments
%%                  for dual anonymous review.
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
%% to improve the accessibility of the compiled pdf file.
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\psr}{J0027$-$1956}
\newcommand{\psrdm}{20.869}
\newcommand{\psrdmerr}{0.005}
\newcommand{\psrnick}{J0036$-$1033}
\newcommand{\dmunit}{pc$\,$cm$^{-3}$}
\newcommand{\Sd}{S$_\rm{d}$}
\newcommand{\nullfraction}{78\%}
\newcommand{\ExB}{\ensuremath{{\bf E}\times{\bf B}}}

@@

{\bf
Pulsars that both show complicated drifting behavior and are bright enough for single pulse analysis, are relatively rare.
}
Even a cursory glance at the drift bands of \psr{} show that its drifting behavior is indeed complex and bright, making it a welcome addition to this category of pulsar.
{\bf 
\psr{} is also found to contain null sequences, a phenomenon which is known to be connected to sub-pulse drifting, originally noticed by \citet{Lyne1983}, and further attested by recent studies of PSRs J1727$-$2739 \citep{Wen2016}, B1819$-$22 \citep{Janagal2022}, and the Vela pulsar \citep{Wen2020}.
}

@@

Follow-up of the initial MWA detection with the full 80 minutes, as well as in other archival MWA observations, revealed that the pulsar has a very large nulling fraction ($\sim$~\nullfraction{})\footnote{The nulling fraction estimate comes only from analysis of the MWA observations.}, with some nulling sequences {\bf (excluding occasional intermittent pulses) potentially exceeding $20\,$min (i.e. the length of the first observation on MJD 59002, which contained \textit{only} a few intermittent pulses, as shown in Fig. \ref{fig:maxima})}.
A full analysis of the nulling properties is presented in Section \S\ref{sec:nulling} below.
Despite the fact that \psr{} was serendipitously discovered in the first ten minutes, it may well be the harbinger for a population of relatively bright, intermittent pulsars which the SMART survey, with its $80$-minute dwell times, is well-suited to detect.
A similar strategy is employed in the Low Frequency Array (LOFAR) Tied-Array All-Sky Survey \citep[LOTAAS;][]{Sanidas2019} and the low latitude ($|b| \le 5^\circ$) segment of the High Time Resolution Universe survey \citep[HTRU;][]{Keith2010}, which use dwell times of approximately one hour.

@@

The other {\bf five observations} were taken when the MWA was in its extended configuration, giving a tied-array beam size of $\sim3^\prime$ at $155\,$MHz.
Localization within {\bf the four observations in which the pulsar was bright enough} followed the identical procedure as used in \citet{Swainston2021}, i.e.\ gridding the candidate positions with overlapping tied-array beams and using signal-to-noise and the estimated beam shape to produce a likelihood map of the source position.
It is known that the ionosphere is capable of shifting the apparent position of sources by an appreciable fraction ($\sim10$-$30\%$) of the tied-array beam, and that these offsets are not necessarily corrected by the calibration process when the calibration observations are separated from the target observation, both spatially and temporally \citep{Swainston2022}.
Nevertheless, the localization afforded by the extended array observations was sufficiently precise (a few arcminutes) to warrant follow-up with a high-resolution imaging telescope.
Accordingly, images made with Band 3 ($300$-$500\,$MHz) data from the upgraded Giant Metrewave Radio Telescope \citep[uGMRT;][]{Gupta2017}, obtained under Director's Discretionary Time (DDTC185), enabled us to confirm the pulsar's position to an accuracy of $\sim1^\prime$.
The right ascension and declination of the final best position, consistent with both the uGMRT imaging and the MWA extended configuration observations, are {\bf 00h26m36s and $-$19$^\circ$55'59''}.

@@

\begin{figure*}[p]
    \centering
    \includegraphics[width=0.48\textwidth]{1226062160_nullhist.png}
    \includegraphics[width=0.48\textwidth]{1274143152_nullhist.png} \\
    \includegraphics[width=0.48\textwidth]{1275094456_nullhist.png}
    \includegraphics[width=0.48\textwidth]{1275172216_nullhist.png} \\
    \includegraphics[width=0.48\textwidth]{1275178816_nullhist.png}
    \includegraphics[width=0.48\textwidth]{1283104232_nullhist.png}
    \caption{{\bf The pulse energy histograms for each of the six observations, for matched-size on- and off-pulse regions of the pulse stacks. The ObsIDs are given above each plot. The on-pulse histograms include all pulses, regardless of whether they were classified as nulls by the method described in the text.}}
    \label{fig:nullhist}
\end{figure*}
